\textbf{ Matching forecast time-scale with ecological process }

### 2019-05-07 doc/manuscript.tex
started adding references

A recent flurry of synthesis papers in high impact journals attests to surging interest in ecological forecasting. Many of the papers 
argue that a commitment to ecological forecasting is necessary to guide the environmental policy and management
decisions that will become increasingly critical under global change (Clark 2001, Mouquet et al. 2017, Dietze et al. 2018).
A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery (Houlahan et al. 2017, Dietz 2017, Dietz et al. 2018).
Taken together, these papers, along with a new textbook (Dietze 2017) represent an emerging consensus that forecasting should be a priority for ecological research.

No such consensus is available for exactly how to tackle this challenge.
Ecologists rely on many different approaches and models for making predictions, and 
frequently argue about the virtues of process-based vs. phenomenological models (refs?).
We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g., Adler et al. 2012),
and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions (refs?).
But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
uncertainty (refs? Tredennick et al. 2017?).

The pragmatic solution to these debates is clear: we should forecast using the models that make the best predictions.
Identifying these models requires rigorously testing predictions against observations. However, many common ecological predictions,
such as a species' geographic range in the year 2100, are virtually impossible to test (but see paleo hindcasting refs,Williams refs? Blois?), 
leading Dietze et al (2017) to call for greater emphasis on near-term forecasting. The assumption is that building 
our capacity to make short-term forecasts will ultimately improve our long-term forecasts.

@@

The second model represents what we call a ``temporal approach.'' We correlate interannual variation in an ecological response with interannual variation in the environment, 
typically at just one site. This approach is often used to study population or vital rate fluctuations as a function of weather, 
or primary production as a function of precipitation (Lauenroth and Sala 1992). The advantages of this approach include is focus on temporal dynamics, meaning it 
can predict rates of change, and validation of the resulting near-term predictions are straightforward. However, models built using this approach typically cover 
a limited spatial extent (but see Kleinhesselink and Adler 2018, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
that could alter the underlying process at longer time scales.

@@

%\newpage
%\renewcommand{\refname}{Literature cited}
%\bibliographystyle{Ecology}
%\bibliography{RemovalRefs}

>>>>

A recent flurry of synthesis papers in high impact journals attests to surging interest in ecological forecasting. Many of the papers 
argue that a commitment to ecological forecasting is necessary to guide the environmental policy and management
decisions that will become increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery 
\citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
Taken together, these papers, along with a new textbook \citep{dietze_ecological_2017} represent an emerging consensus that 
forecasting should be a priority for ecological research.

No such consensus is available for exactly how to tackle this challenge.
Ecologists rely on many different approaches and models for making predictions, and 
frequently argue about the virtues of process-based vs. phenomenological models (refs?).
We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g., Adler et al. 2012),
and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions (refs?).
But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
uncertainty (refs? \cite{tredennick_we_2017}?).

The pragmatic solution to these debates is clear: we should forecast using the models that make the best predictions.
Identifying these models requires rigorously testing predictions against observations. However, many common ecological predictions,
such as a species' geographic range in the year 2100, are virtually impossible to test (but see paleo hindcasting refs,Williams refs? Blois?), 
leading \citep{dietze_iterative_2018} to call for greater emphasis on near-term forecasting. The assumption is that building 
our capacity to make short-term forecasts will ultimately improve our long-term forecasts.

@@

The second model represents what we call a ``temporal approach.'' We correlate interannual variation in an ecological response with interannual variation in the environment, 
typically at just one site. This approach is often used to study population or vital rate fluctuations as a function of weather \citep{dalgleish_climate_2011}, 
or primary production as a function of precipitation \citep{lauenroth_long-term_1992}. The advantages of this approach include is focus on temporal dynamics, meaning it 
can predict rates of change, and validation of the resulting near-term predictions are straightforward. However, models built using this approach typically cover 
a limited spatial extent (but see \cite{kleinhesselink_response_2018}, others), and ignore slower processes, such as evolutionary adaptation or turnover in community composition, 
that could alter the underlying process at longer time scales.

@@

\newpage
\renewcommand{\refname}{Literature cited}
\bibliographystyle{Ecology}
\bibliography{references}
